---
title: "IE5374 Project"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Importing the libraries

```{r}

library(tidyverse)
library(maps)
library(viridis)
library(caret)
library(plyr)
library(data.table)
library(randomForest)

```


Importing the dataset

```{r}

PTR_df <- read_csv("property-tax-report Updated.csv")
head(PTR_df)

```

Columns in the dataset

```{r}

names(PTR_df)

```

Omitting some columns

```{r}

PTR_df <- subset(PTR_df, select = -c(NARRATIVE_LEGAL_LINE1, NARRATIVE_LEGAL_LINE2, NARRATIVE_LEGAL_LINE3, NARRATIVE_LEGAL_LINE4, NARRATIVE_LEGAL_LINE5, FOLIO, ZONING_DISTRICT,LOT, PLAN, BLOCK, DISTRICT_LOT, FROM_CIVIC_NUMBER, TO_CIVIC_NUMBER, TAX_LEVY, NEIGHBOURHOOD_CODE, REPORT_YEAR, LAND_COORDINATE))
head(PTR_df)

```

#####################################################################################################################################################################################

EXPLORATORY DATA ANALYSIS

```{r}

nrow(PTR_df)
ncol(PTR_df)


length(unique(PTR_df$PID))
unique(PTR_df$LEGAL_TYPE)

```

Summary of the dataset

```{r}
summary(PTR_df)
```

Cleaning the dataset (Replacing NULL Values with Mean/Mode values of the column)

```{r}

PTR_df$CURRENT_LAND_VALUE[is.na(PTR_df$CURRENT_LAND_VALUE)]<-mean(PTR_df$CURRENT_LAND_VALUE,na.rm=TRUE)
PTR_df$CURRENT_IMPROVEMENT_VALUE[is.na(PTR_df$CURRENT_IMPROVEMENT_VALUE)]<-mean(PTR_df$CURRENT_IMPROVEMENT_VALUE,na.rm=TRUE)
PTR_df$PREVIOUS_LAND_VALUE[is.na(PTR_df$PREVIOUS_LAND_VALUE)]<-mean(PTR_df$PREVIOUS_LAND_VALUE,na.rm=TRUE)
PTR_df$PREVIOUS_IMPROVEMENT_VALUE[is.na(PTR_df$PREVIOUS_IMPROVEMENT_VALUE)]<-mean(PTR_df$PREVIOUS_IMPROVEMENT_VALUE,na.rm=TRUE)


PTR_df$TAX_ASSESSMENT_YEAR[is.na(PTR_df$TAX_ASSESSMENT_YEAR)]<-mode(PTR_df$TAX_ASSESSMENT_YEAR)
PTR_df$YEAR_BUILT[is.na(PTR_df$YEAR_BUILT)]<-mode(PTR_df$YEAR_BUILT)
PTR_df$BIG_IMPROVEMENT_YEAR[is.na(PTR_df$BIG_IMPROVEMENT_YEAR)]<-mode(PTR_df$BIG_IMPROVEMENT_YEAR)

summary(PTR_df)

```


Current Land Value of each Legal Type

```{r}
plot1 <- data.frame(PTR_df$LEGAL_TYPE, PTR_df$CURRENT_LAND_VALUE)
plot1 <- aggregate(plot1$PTR_df.CURRENT_LAND_VALUE, by = list(Category = plot1$PTR_df.LEGAL_TYPE), FUN=sum)

data <- plot1
data <- data %>% 
  arrange(desc(Category)) %>%
  mutate(prop = x / sum(data$x) *100) %>%
  mutate(ypos = cumsum(prop)- 0.5*prop )



plots1 <- ggplot(data, aes(x="", y=prop, fill=Category)) +
  geom_bar(stat="identity", width=1, color="white") +
  coord_polar("y", start=0) +
  theme_void() + 
  theme(legend.position="none") +
  
  geom_text(aes(y = ypos, label = Category), color = "white", size=6) +
  scale_fill_brewer(palette="Set1")

png('PLOT 1 Current Land Value of each Legal Type.png')
plots1
dev.off()

plots1
```

Current Land Value of each Zone

```{r}

plot2 <- data.frame(PTR_df$ZONING_CLASSIFICATION, PTR_df$CURRENT_LAND_VALUE)
plot2 <- aggregate(plot2$PTR_df.CURRENT_LAND_VALUE, by = list(Category = plot2$PTR_df.ZONING_CLASSIFICATION), FUN=sum)
plot2$id <- seq(1, nrow(plot2))

label_data <- plot2
number_of_bar <- nrow(label_data)
angle <- 90 - 360 * (label_data$id-0.5) /number_of_bar     # I substract 0.5 because the letter must have the angle of the center of the bars. Not extreme right(1) or extreme left (0)

label_data$hjust <- ifelse( angle < -90, 1, 0)
label_data$angle <- ifelse(angle < -90, angle, angle-180)
label_data <- label_data[order(label_data$x),]
label_data$z = c(15,25,35,45,55,65,75,85,95)

colorss = c("blue", "red", "black", "brown", "darkgray", "darkorchid", "indianred4", "tan3", "yellow")

p <- ggplot(label_data, aes(x = as.factor(id), y = z, fill = Category)) +       # Note that id is a factor. If x is numeric, there is some space between the first bar
  geom_bar(stat="identity", fill=alpha(colorss , 0.3)) +
  ylim(-200,400) +
  theme_minimal() +
  theme(
    axis.text = element_blank(),
    axis.title = element_blank(),
    panel.grid = element_blank(),
    plot.margin = unit(rep(-2,12), "cm") 
  ) +
  coord_polar(start = 0) + 
  geom_text(data=label_data, aes(x=id, y = z+10, label=Category, hjust = hjust), color="black", fontface="bold",alpha=3, size=3, angle= 0, inherit.aes = FALSE ) 

png('PLOT 2 Current Land Value of each Zone.png')
p
dev.off()

p


```



Relation Between Current Land Value and Previous Land Value for each Zone

```{r}

plot3 <- subset(PTR_df, select = c(ZONING_CLASSIFICATION, CURRENT_LAND_VALUE, PREVIOUS_LAND_VALUE))

q = aggregate(c(plot3$CURRENT_LAND_VALUE ), by = list(plot3$ZONING_CLASSIFICATION), max)
w = aggregate(c(plot3$PREVIOUS_LAND_VALUE), by = list(plot3$ZONING_CLASSIFICATION), max)

rm(plot3)

plot3 <- data.frame(ZONING_CLASSIFICATION = q$Group.1, A_PREVIOUS_LAND_VALUE = w$x, B_CURRENT_LAND_VALUE = q$x)
plot3 <- plot3 %>% gather("Current_and_Previous", "Value", 2:3)
plot3

p3 <- plot3 %>% ggplot(aes(x = Current_and_Previous, y = Value, group = ZONING_CLASSIFICATION, color = ZONING_CLASSIFICATION)) + geom_line() + theme_classic()


png('PLOT 3 Relation Between Current Land Value and Previous Land Value for each Zone.png')
p3
dev.off()

p3


```

Relation Between Current Improvement Value and Previous Improvement Value

```{r}

plot4 <- subset(PTR_df, select = c(CURRENT_IMPROVEMENT_VALUE, PREVIOUS_IMPROVEMENT_VALUE))
p4 <- ggplot(plot4, aes(x = CURRENT_IMPROVEMENT_VALUE, y = PREVIOUS_IMPROVEMENT_VALUE)) + geom_point() + geom_smooth(method = lm, color = "red", fill = "#69b3a2", se = TRUE) + theme_classic()

png('PLOT 4 Relation Between Current Land Value and Previous Land Value for each Property.png')
p4
dev.off()

p4

```

Relation between Current Land Value and Previous Land Value

```{r}

plot5 <- subset(PTR_df, select = c(CURRENT_LAND_VALUE, PREVIOUS_LAND_VALUE))
p5 <- ggplot(plot5, aes(x = CURRENT_LAND_VALUE, y = PREVIOUS_LAND_VALUE)) + geom_point() + geom_smooth(method = lm, color = "blue", fill = "#69b3a2", se = TRUE) + theme_classic()


png('PLOT 5 Relation Between Current Land Value and Previous Land Value for each Zone.png')
p5
dev.off()

p5

```

Top 10 Most Expensive Streets 

```{r}

plot6 <- subset(PTR_df, select = c(STREET_NAME, CURRENT_LAND_VALUE))  
plot6 <- plot6[order(-plot6$CURRENT_LAND_VALUE),][1:10,]
plot6

p6 <- ggplot(plot6, aes(x = STREET_NAME, y = CURRENT_LAND_VALUE, fill = STREET_NAME)) + geom_bar(stat = "identity") + coord_flip() + theme_classic()

png('PLOT 6 Top 10 Expensive Streets of Vancouver.png')
p6
dev.off()

p6

```

Top 10 Most Expensive Properties in Vancouver


```{r}

plot7 <- subset(PTR_df, select = c(PID, CURRENT_LAND_VALUE))  
plot7 <- plot7[order(-plot6$CURRENT_LAND_VALUE),][1:10,]
plot7

p7 <- ggplot(plot7, aes(x=PID, y=CURRENT_LAND_VALUE)) +
  geom_segment( aes(x=PID, xend=PID, y=0, yend=CURRENT_LAND_VALUE), color="grey") +
  geom_point( color="orange", size=4) +
  theme_light() +
  theme(
    panel.grid.major.x = element_blank(),
    panel.border = element_blank(),
    axis.ticks.x = element_blank()
  ) +
  xlab("PID") +
  ylab("CURRENT_LAND_VALUE") + coord_flip()

png('PLOT 7 Top 10 Most expensive Properties of Vancouver.png')
p7
dev.off()

p7

```



Number of properties built in different years

```{r}
plot8 <- count(PTR_df, "YEAR_BUILT")
plot8

setnames(plot8, old = c('freq'), new = c('count'))

data <- plot8

# Compute percentages
data$fraction = data$count / sum(data$count)

# Compute the cumulative percentages (top of each rectangle)
data$ymax = cumsum(data$fraction)

# Compute the bottom of each rectangle
data$ymin = c(0, head(data$ymax, n=-1))
 
# Make the plot
p8 <- ggplot(data, aes(ymax=ymax, ymin=ymin, xmax=4, xmin=3, fill=YEAR_BUILT)) +
     geom_rect() +
     coord_polar(theta="y") + # Try to remove that to understand how the chart is built initially
     xlim(c(2, 4)) # Try to remove that to see how to make a pie chart

png('PLOT 8 Recency of Properties of Vancouver.png')
p8
dev.off()

p8



```


#####################################################################################################################################################################################

MULTIPLE LINEAR REGRESSION

Multiple Linear  Regression to learn the pattern between previous land value and current land value to predict the future land value (Value of the land in 5 years) using features PID, zone classification, Street name, previous land value, current land value, year built.


Using the Previous land value, previous improvement value, year built to train and predict the current land value of the property.
Using the same model to predict the future land value.

Big improvement was not contributing to the model

Building the model

```{r}

PTR_df$YEAR_BUILT <- as.numeric(PTR_df$YEAR_BUILT)
PTR_df$BIG_IMPROVEMENT_YEAR <- as.numeric(PTR_df$BIG_IMPROVEMENT_YEAR)

df_for_pred <- subset(PTR_df, select = c(PREVIOUS_LAND_VALUE, PREVIOUS_IMPROVEMENT_VALUE, YEAR_BUILT, CURRENT_LAND_VALUE))[10:91,]
df_for_pred <- na.omit(df_for_pred)

df_for_pred$PID <- c(1:64)
df_CURRENT_LAND_VALUE <- subset(df_for_pred, select = CURRENT_LAND_VALUE)

model <- lm(df_for_pred$CURRENT_LAND_VALUE ~ df_for_pred$PID + df_for_pred$PREVIOUS_IMPROVEMENT_VALUE + df_for_pred$PREVIOUS_IMPROVEMENT_VALUE + df_for_pred$YEAR_BUILT)

cor(df_for_pred)

df_test <- df_for_pred$CURRENT_LAND_VALUE

df_for_pred <- subset(df_for_pred, select = -CURRENT_LAND_VALUE)
df_for_pred$prediction <- model %>% predict(df_for_pred)

head(df_for_pred$prediction)
head(df_test)

```

Summary of the model

```{r}

summary(model)

```



Using Random Forest to train our model

```{r}

df_rf_pred <- subset(PTR_df, select = c(PREVIOUS_LAND_VALUE, PREVIOUS_IMPROVEMENT_VALUE, YEAR_BUILT, CURRENT_LAND_VALUE))[10:91,]
df_rf_pred <- na.omit(df_rf_pred)
df_rf_pred

df_rf_pred$PID <- c(1:64)
df_CURRENT_LAND_VALUE_rf <- subset(df_rf_pred, select = CURRENT_LAND_VALUE)

rf <- randomForest(df_rf_pred$CURRENT_LAND_VALUE ~ df_rf_pred$PID + df_rf_pred$PREVIOUS_IMPROVEMENT_VALUE + df_rf_pred$PREVIOUS_IMPROVEMENT_VALUE + df_rf_pred$YEAR_BUILT, proximity = TRUE)

cor(df_rf_pred)

df_test_rf <- df_rf_pred$CURRENT_LAND_VALUE

df_rf_pred <- subset(df_rf_pred, select = -CURRENT_LAND_VALUE)

df_rf_pred$prediction <- rf %>% predict(df_rf_pred)

head(df_rf_pred$prediction)
head(df_test_rf)

testing <- data.frame(PID = df_rf_pred$PID, CURRENT_LAND_VALUE = df_test_rf)

R2(df_test_rf, df_rf_pred$prediction)

```
Graph for future land Value 

```{r}

df_rf_pred_1 <- subset(PTR_df, select = c(CURRENT_LAND_VALUE, CURRENT_IMPROVEMENT_VALUE, YEAR_BUILT))[10:91,]
df_rf_pred_1 <- na.omit(df_rf_pred_1)

names(df_rf_pred_1)[1] ="PREVIOUS_LAND_VALUE"
names(df_rf_pred_1)[2] ="PREVIOUS_IMPROVEMENT_VALUE"

df_rf_pred_1$prediction <- rf %>% predict(df_rf_pred_1)
df_rf_pred_1$PID <- c(1:64)
df_rf_pred_1 <- subset(df_rf_pred_1, select = c(prediction, PID, PREVIOUS_LAND_VALUE))

ggplot() + geom_line(data = testing, aes(x = PID, y = CURRENT_LAND_VALUE), color = "blue") + geom_line(data = df_rf_pred_1, aes(x = PID, y = prediction), color = "red") + theme_classic() + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) 

```

```{r}

df_rf_pred_improvement <- subset(PTR_df, select = c(PREVIOUS_IMPROVEMENT_VALUE, BIG_IMPROVEMENT_YEAR, YEAR_BUILT, CURRENT_IMPROVEMENT_VALUE))[10:91,]
df_rf_pred_improvement <- na.omit(df_rf_pred_improvement)
df_rf_pred_improvement

rf_1 <- randomForest(df_rf_pred_improvement$CURRENT_IMPROVEMENT_VALUE ~ df_rf_pred_improvement$PREVIOUS_IMPROVEMENT_VALUE + df_rf_pred_improvement$BIG_IMPROVEMENT_YEAR + df_rf_pred_improvement$YEAR_BUILT, proximity = TRUE)

cor(df_rf_pred_improvement)

df_rf_pred_improvement_current <- df_rf_pred_improvement$CURRENT_IMPROVEMENT_VALUE

df_rf_pred_improvement <- subset(df_rf_pred_improvement, select = -CURRENT_IMPROVEMENT_VALUE)
df_rf_pred_improvement

df_rf_pred_improvement$prediction <- rf_1 %>% predict(df_rf_pred_improvement)
df_rf_pred_improvement

head(df_rf_pred_improvement$prediction)
head(df_rf_pred_improvement_current)

testing <- data.frame(Current_improvement_Value = df_rf_pred_improvement_current, Prediction = df_rf_pred_improvement$prediction)
testing

R2(testing$Current_improvement_Value, testing$Prediction)


```

Graph for future Improvement Value 

```{r}

testing$PID <- c(1:64)

df_rf_pred_improvement_1 <- subset(PTR_df, select = c(BIG_IMPROVEMENT_YEAR, YEAR_BUILT, CURRENT_IMPROVEMENT_VALUE))[10:91,]
df_rf_pred_improvement_1 <- na.omit(df_rf_pred_improvement_1)
names(df_rf_pred_improvement_1)[3] ="PREVIOUS_IMPROVEMENT_VALUE"

df_rf_pred_improvement_1$prediction <- rf_1 %>% predict(df_rf_pred_improvement_1)
df_rf_pred_improvement_1$PID <- c(1:64)
df_rf_pred_improvement_1 <- subset(df_rf_pred_improvement_1, select = c(prediction, PID, PREVIOUS_IMPROVEMENT_VALUE))


ggplot() + geom_line(data = testing, aes(x = PID, y = Current_improvement_Value), color = "blue") + geom_line(data = df_rf_pred_improvement_1, aes(x = PID, y = prediction), color = "red") + theme_classic() + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) 

```


#####################################################################################################################################################################################

